---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

###Set time format and working directory

```{r, results="hide", echo=TRUE}
setwd("c:\\WorkDir\\Study\\Coursera\\ReproduciableResearch\\")
Sys.setlocale("LC_TIME", "English")
```

###Load data

```{r, echo=TRUE}
data<-read.csv("c:\\WorkDir\\Study\\Coursera\\ReproduciableResearch\\repdatadataactivity\\activity.csv")
```

###Process/transform the data

```{r,echo=TRUE}
data$date<-factor(data$date)
data$timeInterval<-(data$interval-data$interval%%100)/100*60+data$interval%%100
```

## What is mean total number of steps taken per day?
###Calculate the total number of steps taken per day
```{r, echo=TRUE}
stepDay<-sapply(split(data, data$date), function(x) sum(x$steps))
stepDay
```

###Compare histogram and barplot 
```{r, echo=TRUE}
hist(stepDay,col="blue",breaks=20)
barplot(stepDay)
```

###Calculate and report the mean and median of the total number of steps taken per day
```{r, echo=TRUE}
mean(stepDay,na.rm=TRUE)
median(stepDay,na.rm=TRUE)
```


## What is the average daily activity pattern?
###Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r, echo=TRUE}
stepInterval<-sapply(split(data, data$timeInterval), function(x) mean(x$steps,na.rm=TRUE))
plot(names(stepInterval),stepInterval, type="l",xlab="Minutes past from 0:00")
```


###Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r, echo=TRUE}
ans<-names(stepInterval[stepInterval==max(stepInterval)])
paste(as.character(as.integer(as.integer(ans)/60)), as.character(as.integer(ans)%%60),sep=":")
```



## Imputing missing values
###Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r, echo=TRUE}
summary(data$steps)["NA's"]
```


###Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

Missing values will be replaced with an average value of steps taken in the corresponding 5-min interval in days when such values are available. Moreover if the average number is not integer we transform it to integer using function as.integer().

###Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r, echo=TRUE}
datanoNA<-data
for(var in which(is.na(data$steps))) {datanoNA$steps[var]<-as.integer(stepInterval[as.character(data$timeInterval[var])])}
```


###Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?
```{r, echo=TRUE}
stepsInDayNoNA<-sapply(split(datanoNA, datanoNA$date), function(x) sum(x$steps))
stepsInDayNoNA
mean(stepsInDayNoNA,na.rm=FALSE)
median(stepsInDayNoNA,na.rm=FALSE)
sum(stepsInDayNoNA)
sum(stepDay,na.rm=TRUE)
hist(stepsInDayNoNA,col="blue",breaks=20)
stepsInDayNoNA-stepDay
```



## Are there differences in activity patterns between weekdays and weekends?

###Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r, echo=TRUE}
wEnd<-function(x){
if (x=="Sunday" | x=="Saturday")
{ans="weekend"}
else {ans="weekday"}
ans
}
datanoNA$weekend<-factor(sapply(weekdays(strptime( as.character(datanoNA$date),"%Y-%m-%d" )),wEnd))
```


###Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.


```{r, echo=TRUE}
avStep<-sapply(split(datanoNA, datanoNA$weekend), function(x) {sapply(split(x,x$timeInterval), function(y) mean(y$steps) )} ) 
datanoNA$IntervalMeanSteps<-replicate(dim(datanoNA)[1],0)
for(i in 1:dim(datanoNA)[1])
{
if (datanoNA$weekend[i]=="weekday") { datanoNA$IntervalMeanSteps[i]<-avStep[ datanoNA$timeInterval[i]/5+1,1]}
else { datanoNA$IntervalMeanSteps[i]<-avStep[datanoNA$timeInterval[i]/5+1,2]} 
}
library(lattice)
xyplot( IntervalMeanSteps  ~ interval  | weekend,data=datanoNA,layout=c(1,2), type="l")
